Generation of TypeScript Declaration Files from JavaScript Code

### 2020-09-10 programming-2021/main.tex

This work presents \texttt{dts-generate}, a tool that generates
TypeScript declaration files for JavaScript libraries uploaded to the \NPM{}
registry. It extracts code examples from the documentation written by
the developer, executes the library driven by the examples, gathers
run-time information, and generates a declaration file based on this
information. To evaluate the tool, 244 declaration files were
generated directly from an \NPM{} module and 82 of these were
compared with the corresponding declaration file provided on
DefinitelyTyped. We obtained positive results for 52 files, which
exhibited no differences at all or differences that can be resolved by
modifying the developer-provided examples. 
\end{abstract}

@@

Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
for 244 modules out of 6029 modules and we identified 60 modules that have only the
features implemented by \texttt{dts-generate}. We obtained positive results for 52 of
them. Section~\ref{sec:experiments-evaluation} provides a detailed explanation of the overall
quality of the generated files.
Section~\ref{sec:experiments-declaration-files-generation} presents examples of the generated declaration files for
templates \textbf{module}, \textbf{module-class}, and \textbf{module-function}.

@@

\subsection{Evaluation}
\label{sec:experiments-evaluation}
We analyzed 60 of the 244 generated files. The remaining 184 files contained at least one
of the unimplemented TypeScript features so that they could not be considered. These
features were identified by \texttt{dts-parse} as tags.

As shown in \figref{fig:experiments-results-module-is-uuid}, the code examples have a
direct influence on the quality of the generated declaration file. We introduced the
concept of \texttt{solvable} and \texttt{unsolvable} differences in
\secref{sec:dts-compare}. Modules containing only \texttt{solvable} differences were
considered as positive results. We identified 52 such modules. We manually completed the
examples for 10 of those modules. Applying \texttt{dts-generate} with the completed
examples generates declaration files, which are equivalent to the files from DefinitelyTyped, as
shown in \figref{fig:experiments-results-manually-completed-examples}.

@@

We discovered only 8 modules with \texttt{unsolvable} differences. Most of them were
caused by the same error: 7 modules incorrectly declared an interface for the
\texttt{string} type. This interface contains standard properties and methods of the
JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}.

>>>>

This work presents \texttt{dts-generate}, a tool that generates
TypeScript declaration files for JavaScript libraries uploaded to the \NPM{}
registry. It extracts code examples from the documentation written by
the developer, executes the library driven by the examples, gathers
run-time information, and generates a declaration file based on this
information. To evaluate the tool, 244 declaration files were
generated directly from an \NPM{} module and 70 of these were
compared with the corresponding declaration file provided on
DefinitelyTyped. We obtained positive results for all of these files, which
exhibited no differences at all or differences that can be resolved by
modifying the developer-provided examples. 
\end{abstract}

@@

Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
for 244 modules out of 6029 modules and we identified 70 modules that have only the
features implemented by \texttt{dts-generate}. We obtained positive results for all of
them. Section~\ref{sec:experiments-evaluation} provides a detailed explanation of the overall
quality of the generated files.
Section~\ref{sec:experiments-declaration-files-generation} presents examples of the generated declaration files for
templates \textbf{module}, \textbf{module-class}, and \textbf{module-function}.

@@

\subsection{Evaluation}
\label{sec:experiments-evaluation}
We analyzed 70 of the 244 generated files. The remaining 174 files contained at least one
of the unimplemented TypeScript features so that they could not be considered. These
features were identified by \texttt{dts-parse} as tags.

As shown in \figref{fig:experiments-results-module-is-uuid}, the code examples have a
direct influence on the quality of the generated declaration file. We introduced the
concept of \texttt{solvable} and \texttt{unsolvable} differences in
\secref{sec:dts-compare}. Modules containing only \texttt{solvable} differences were
considered as positive results. All of the analyzed files contained only \texttt{solvable} differences. We manually completed the
examples for 10 of those modules. Applying \texttt{dts-generate} with the completed
examples generates declaration files, which are equivalent to the files from DefinitelyTyped, as
shown in \figref{fig:experiments-results-manually-completed-examples}.

@@
